\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-07-04 eg6-letters/NIM/EG6_RTPC_NIM.tex
NIM:  z-resolutions now after subtracting CLAS contribution
rm Line numbers
Add Stepan's comments

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\linenumbers
\begin{document}

@@

Until recently, the Thomas Jefferson National Accelerator Facility, in 
Newport News, Virginia, USA, has provided high power electron beams of 
up to 6 GeV energy and 100$\%$ duty factor to three experimental Halls (A, B, C) 
simultaneously. The CLAS spectrometer~\cite{CLASref}, located
in Hall-B, is based on a superconducting toroidal magnet and composed of 
several sub-detectors. Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer:
\begin{itemize}
 \item Three regions of Drift Chambers (DC) for the tracking of charged 
       particles~\cite{DCref}.
 \item Superconducting toroidal magnet to bend the trajectories 
       of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item Threshold Cherenkov Counters for electron identification at momenta 
       $<2.7$ GeV~\cite{CCref}.
 \item Scintillation Counters to identify charged hadrons by measuring their 
       time of flight~\cite{TOFref}.
 \item Electromagnetic Calorimeters (EC) for identification of electrons, 
       photons and neutrons~\cite{ECref}.
\end{itemize}

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a nucleon or nuclei) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005~\cite{FX}.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$~\cite{Hyon-suk}. To protect the 
CLAS detector and the IC from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the detectors. 
To detect recoiling $\alpha$ particles from the coherent DVCS on helium, a new 
radial time projection chamber (RTPC) was developed to track low energy nuclear 
fragments. The solenoid field was used to track and measure momentum of 
particles in RTPC. This 
setup was used during a three months experimental run~\cite{proposal1,proposal2}
in 2009 with a longitudinally polarized electron beam of 130~nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.

The original design of the RTPC~\cite{BONUS-NIM} was developed for the BoNuS 
experiment at Jefferson Lab which took data in CLAS in 2005. An enhanced design, 
used in the 2009 DVCS experiment, is described in section \ref{sec_design} 
of this paper. Significant improvements were made to the RTPC mechanical 
structure and fabrication technique that both increased the acceptance and 
reduced the amount of material in the path of the outgoing particles. The 
data acquisition system described in section \ref{sec_readout} was reconfigured 
to increase the event readout rate. The calibration methods 
are discussed in section \ref{sec_calib}. Finally, the overall performance
of the RTPC is described in section \ref{sec_perfor}.

@@

We detail here the different regions shown in Fig.~\ref{fig:RTPC_1_4} starting 
from the beam line towards larger radius:\\
\begin{itemize}
   \item The 6 atm helium-4 target extends along the beam line axis; it 
      is a 284~mm long, 6~mm diameter straw with a 27-$\mu$m Kapton wall.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gap region extends between 20~mm and 30~mm and is filled with the 
      gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar set at 4260~V 
      to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away from the beam axis. The average 
      electric field in this region is perpendicular to the beam and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower Voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The drift gas used in the experiment is a 80-20\% Ne-DME mixture. This choice 
has been made in order to balance the energy deposit, which is critical
for proper particle identification, with a reasonable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the axial 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this gas mixture. This is already a significant angle,
and heavier gas have a tendency to give larger angle, increasing significantly
the length of the drift paths as the electrons would 
make a long spiral around the beam line to reach the GEMs.

One of the important challenge in developing the radial TPC was to obtain a 
good support structure for the GEM foils to allow a tractable 
installation of the GEMs. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each end of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
gluing and soldering operations. The rigidity of the GEM foils was enough
for the structure to be self-supporting and only the upstream end of the cylinder 
was fixed to the main mechanical structure. This design only left a 
light fiberglass ring in the downstream end, reducing to a minimum secondary 
interactions.

@@

The RTPC electron collection system has 3200 readout pads. These elements are
parallel to the GEMs, 69 mm from the central axis.
Figure~\ref{fig:PADs} shows a schematic drawing of the size and 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width. The shift between the rows permits the reduction of aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in Fig.~\ref{fig:PADs} 
shows how pads are grouped to 16 channel pre-amplifiers. The pre-amplifier boards, 
already employed in the BoNuS RTPC~\cite{BONUS-NIM}, serve the dual purpose of 
inverting the RTPC signals polarity -- from negative to positive -- to match the 
requirements of the subsequent readout system, and drive the 6~m long ribbon 
cable that connects to it.

@@

The readout system is based on the Front End Electronic (FEE) boards originally 
developed for the ALICE TPC readout system at CERN~\cite{ALICE-FEE}. Each 
readout channel is made of three main components: a charge sensitive, low 
impedance shaping amplifier, a 10 bit, 25-MHz digitizer, and a digital circuit 
that implements online-processing algorithms (pedestal subtraction, zero 
suppression, and tail cancellation). The amplifier is implemented in a 
fully-analogue ASIC, PASA~\cite{ALICE-PASA}, while the signal 
digitization and elaboration part is implemented in the digital ASIC, 
ALTRO~\cite{ALICE-ALTRO}. Both ASICs host 16 channels. The full readout 
chain is implemented on a single Front End Card (FEC), hosting 128 channels. 
FECs are hosted in custom crates, mounted close to the detector. A readout 
control unit (RCU) board is used to distribute the trigger signal to FECs 
and for data readout. Each RCU can handle up to 25 FECs with communication 
between them performed through a 
custom back-plane, implementing a low-Voltage signal bus. The RCU communicates 
with the CLAS data acquisition (DAQ) system 
trough a 200 MB/s optical link, connected to a data acquisition PC hosting 
a ReadOut Receiver Card (RORC) used to interface with 
the CLAS DAQ system. An Ethernet link is also present, for slow-controls and 
monitoring.

@@

To measure the drift speed, we use the time profile of all hits in the 
chamber shown in Fig.~\ref{fig:TDC_profile}. We can clearly observe the 
dropping edge expected from geometrical considerations. We define a value $T_{Max/2}$ at 
which the dropping edge passes half the maximum number of hits in the 
histogram. This value was measured in bins along the 200 mm RTPC's length to take into 
account variations in the electric and magnetic field in the RTPC (see 
Fig.~\ref{fig:RunNumber_61551_TDCmax_Zslice}).

@@

Due to the non perfect experimental conditions, in particular variations in the 
gas mixture\footnote{It was noted during the experiment that leaks appeared,
most likely in one of the $4 \mu$m foils between the beam and drift regions,
and the gas flow had to be increased in few occasions to compensate for this.}, 
the drift speed changes during the three months long experimental run.  
Figure~\ref{fig:Drift_run_number_1} shows the $T_{Max/2}$ values for 
individual runs (approximately 2 hours long). We observe a significant 
variation of the drift speed over time that need to be accounted for 
in track reconstruction.

@@

In summary, we obtain from our calibration a parametrization of the drift speed 
as a function of both position along the beam axis and the run number. These 
functions were extracted for our entire data set and implemented in the 
track reconstructions code.

@@

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. Software exists to calculate them, in particular
MAGBOLTZ~\cite{MAGBOLTZ}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used such calculation as a 
first calibration, but, 
as can be seen with the drift speed, gas composition is far from stable in the 
chamber. Moreover, the 4 $\mu$m foil used as a cathode is easily deformed, such
that we expect the geometry accuracy to be of few millimeters, directly impacting 
our knowledge of the electric field. These problems, already 
encountered for the BoNuS RTPC calibration~\cite{BONUS-NIM}, motivated the 
acquisition of specific calibration runs. These were taken with a lower energy 
electron beam (1.204 and 1.269 GeV) to enhance the cross section of the elastic 
scattering ($e^{4}$He$\rightarrow e^{4}$He). In this process, the measurement of
the electron kinematic allows to calculate directly the kinematic of the Helium nucleus. 
By comparing the alpha particle's calculated momentum and angle of recoil to the 
measurement in the RTPC, we tuned the drift paths independently of our 
knowledge of the exact chamber's conditions.

Based on the kinematic of electrons in the calibration data, 
we generate the helium nucleus in our RTPC GEANT4 simulation~\cite{GEANT4}. Then 
we compare the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber. 
Because of the magnetic field, the drift paths are not linear in the RTPC. So 
to perform the extraction, we make a first approximation with a linear 
dependence between the radius of emission and the time of detection, and then 
refine our result. As it happens, the curvature is minimal and this process 
converges already on the second iteration.

@@

To calibrate the gains, we compare the experimental ADCs to the energy deposited 
for each pad individually in GEANT4 by similar simulated tracks (using the 
same elastic events as for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the DAQ features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as shown in 
Fig.~\ref{fig:EVENT_adc_tdc}. The gain for each pad is calculated 
as the ratio of the measured ADCs to the simulated deposited energy.  
Then, these gains are refined using correcting factors obtained from a 
sample of good tracks. For each track, we calculate the corrected 
energy deposit on a pad and compare it to the average
deposit recorded by the other pads. This last step was implemented to 
avoid any bias from this simulation based calibration method.
Final results are shown in Fig.~\ref{fig:dedx_p_exp_2nd}.

@@

\section{Track Reconstruction}

@@

The first is an oscillatory noise located early in the readout time window, 
shown in the top panel of Fig.~\ref{fig:noise} for a particularly noisy 
channel. Its amplitude is not dissimilar to those of real tracks, and about 
18\% of the readout channels exhibit large contributions from this noise 
characteristic. Due to its unique time-energy correlation for the given 
channels, the noise could be removed on an event by event and channel by 
channel basis without significant loss of good signals. The result of the
procedure is illustrated in the bottom panel of Fig.~\ref{fig:noise}.

The second is a coherent noise affecting about 25\% of the preamplifiers, where 
the signature is simultaneous hits in most of the 16 channels in a preamp 
group.   An event-based technique to identify and remove it was developed based 
on counting simultaneous hits in preamp group, and, if sufficiently large, 
perform a dynamic pedestal subtraction based on the average ADC of the 
channels' neighbors within the preamp group.

@@

\subsection{Resolution}
Elastic scattering was used to estimate the tracking resolution of the RTPC 
based on the residual between the expected and measured $^4$He tracks.  The RTPC
resolutions, after removing contributions from the electron, are shown in Table
~\ref{tab:reso}, and are very similar for the two halves of the RTPC.  Note, the
$\theta$- and $z$-resolutions are highly correlated.
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|cccc|}
  \hline
& $\sigma_{z}$ &  $\sigma_{\theta}$ & $\sigma_{\phi}$ & $\sigma_{p}$\\
\hline
Left &  6.0 mm & 3.8$^{\circ}$ & 1.9$^{\circ}$ & 9$\%$ \\
Right & 7.4 mm & 4.0$^{\circ}$ & 1.9$^{\circ}$ & 8$\%$\\
\hline
\end{tabular}
\caption{The resolutions of the two modules of the RTPC for $z$-vertex, polar and azimuthal angles, and momentum.}
\label{tab:reso}
\end{center}
\end{table}

\subsection{Efficiency}
The previous discussions have shown that the two modules of the RTPC have 
slightly different behavior and give also different particle yield. However, 
this should not necessarily be linked to different efficiencies within the RTPC. 
Indeed, there is a complicated convolution of CLAS and the RTPC acceptances 
that can significantly affect the yields and kinematics of the detected 
particles. To clarify this question, we measured the efficiency of the RTPC
using elastic scattering on $^4$He by comparing the inclusive yield, based
only on electron detection, to the exclusive elastic yield, where the Helium
recoil is also detected (see Fig.~\ref{fig:w}). We present in Fig.~\ref{fig:rtpc_eff}
the results for the two modules of the detector. We observe that the left and 
the right modules have similar efficiencies except near the target windows.

>>>>

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\begin{document}

@@

Until recently, the Thomas Jefferson National Accelerator Facility, in 
Newport News, Virginia, USA, has provided high power electron beams of 
up to 6 GeV energy and 100$\%$ duty factor to three experimental Halls (A, B, C) 
simultaneously. The CLAS spectrometer~\cite{CLASref}, located
in Hall-B, was based on a superconducting toroidal magnet and composed of 
several sub-detectors. Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer:
\begin{itemize}
 \item Three regions of Drift Chambers (DC) for the tracking of charged 
       particles~\cite{DCref}.
 \item Superconducting toroidal magnet to bend the trajectories 
       of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item Threshold Cherenkov Counters (CC) for electron identification at momenta 
       $<2.7$ GeV~\cite{CCref}.
 \item Scintillation Counters (SC) to identify charged hadrons by measuring their 
       time of flight~\cite{TOFref}.
 \item Electromagnetic Calorimeters (EC) for identification of electrons, 
       photons and neutrons~\cite{ECref}.
\end{itemize}

@@

For certain experiments the base CLAS system was complimented with ancillary 
detectors. For exemple, the measurement of the Deeply Virtual Compton 
Scattering (DVCS) process ($eH \rightarrow e' H' \gamma$, where $H$ is a 
nucleon or nuclei) necessitates an upgrade of the photon detection system. 
Indeed, with a 6 GeV electron beam, the majority of DVCS photons are produced 
at very forward angles, where the acceptance of the EC was poor. To extend the 
detection range, an inner calorimeter (IC) was built for the E01-113 
experiment in 2005~\cite{FX}. The IC is constructed from 424 lead-tungstate 
(PbWO$_{4}$) crystals, covering polar angles between 5$^{\circ}$ and 
15$^{\circ}$~\cite{Hyon-suk}. To protect the 
CLAS detector and the IC from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the detectors. 
To detect recoiling $\alpha$ particles from the coherent DVCS on helium, a new 
radial time projection chamber (RTPC) was developed to track low energy nuclear 
fragments. The solenoid field was used to track and measure momentum of 
particles in RTPC. The CLAS detector supplemented with IC and RTPC was used 
during a three months experimental run~\cite{proposal1,proposal2}
in 2009 with a longitudinally polarized electron beam of 130~nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.

The original design of the RTPC was developed for the BoNuS 
experiment at Jefferson Lab which took data in CLAS in 2005~\cite{BONUS-NIM}. An enhanced design, 
used in the 2009 DVCS experiment, is described in section \ref{sec_design} 
of this paper. Significant improvements were made to the RTPC mechanical 
structure and fabrication technique that both increased the acceptance and 
reduced the amount of material in the path of the outgoing particles. The 
data acquisition system described in section \ref{sec_readout} was reconfigured 
to increase the event readout rate. The tracking algorithm is discussed in section
\ref{sec_tracking}. The calibration methods 
are discussed in section \ref{sec_calib}. Finally, the overall performance
of the RTPC is described in section \ref{sec_perfor}.

@@

We detail here the different regions shown in Fig.~\ref{fig:RTPC_1_4} starting 
from the beam line towards larger radius:\\
\begin{itemize}
   \item The 6 atm helium-4 target extends along the beam line axis; it 
      is a 284~mm long, 6~mm diameter straw with a 27-$\mu$m Kapton wall.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gap region extends between 20~mm and 30~mm and is filled with the 
      gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar set at $- 4260$~V 
      to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away from the beam axis. The average 
      electric field in this region is perpendicular to the beam and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The drift gas used in the experiment is a 80-20\% Ne-DME mixture. This choice 
has been made in order to balance the energy deposit, which is critical
for proper particle identification, with a reasonable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the axial 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this gas mixture.

One of the important challenge in developing the radial TPC was to obtain a 
good support structure for the GEM foils to allow a tractable 
installation of the GEMs. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each end of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
gluing and soldering operations. The rigidity of the GEM foils was enough
for the structure to be self-supporting and only the upstream end of the cylinder 
was fixed to the main mechanical support structure. This design only left a 
light fiberglass ring in the downstream end, reducing to a minimum secondary 
interactions.

@@

The RTPC electron collection system has 3200 readout pads. These elements are
parallel to the GEMs, 69 mm from the central axis.
Figure~\ref{fig:PADs} shows a schematic drawing of the size and 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width. The shift between the rows permits the reduction of aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in Fig.~\ref{fig:PADs} 
shows how pads were grouped on 16 channel pre-amplifiers. The pre-amplifier boards, 
already employed in the BoNuS RTPC~\cite{BONUS-NIM}, serve the dual purpose of 
inverting the RTPC signals polarity -- from negative to positive -- to match the 
requirements of the subsequent readout system, and drive the 6~m long ribbon 
cable that connects to it.

@@

The readout system is based on the Front End Electronic (FEE) boards originally 
developed for the ALICE TPC readout system at CERN~\cite{ALICE-FEE}. Each 
readout channel is made of three main components: a charge sensitive, low 
impedance shaping amplifier, a 10 bit, 25-MHz digitizer, and a digital circuit 
that implements online-processing algorithms (pedestal subtraction, zero 
suppression, and tail cancellation). The amplifier is implemented in a 
fully-analogue ASIC, PASA~\cite{ALICE-PASA}, while the signal 
digitization and elaboration part is implemented in the digital ASIC, 
ALTRO~\cite{ALICE-ALTRO}. Both ASICs host 16 channels. The full readout 
chain is implemented on a single Front End Card (FEC), hosting 128 channels. 
FECs are hosted in custom crates, mounted close to the detector. A readout 
control unit (RCU) board is used to distribute the trigger signal to FECs 
and for data readout. Each RCU can handle up to 25 FECs with communication 
between them performed through a 
custom back-plane, implementing a low-voltage signal bus. The RCU communicates 
with the CLAS data acquisition (DAQ) system 
trough a 200 MB/s optical link, connected to a data acquisition PC hosting 
a ReadOut Receiver Card (RORC) used to interface with 
the CLAS DAQ system. An Ethernet link is also present, for slow-controls and 
monitoring.

@@

To measure the drift speed, we use the time profile of all hits in the 
chamber shown in Fig.~\ref{fig:TDC_profile}. We can clearly observe the 
dropping edge expected from geometrical considerations. We define a value $T_{Max/2}$ at 
which the dropping edge passes half the maximum number of hits in the 
histogram. This value was measured in bins along the 200~mm RTPC's length to take into 
account variations in the electric and magnetic field in the RTPC (see 
Fig.~\ref{fig:RunNumber_61551_TDCmax_Zslice}).

@@

Due to the non perfect experimental conditions, in particular possible contaminations 
of our gas mixture, the drift speed changed during the three months long experimental 
run\footnote{We had to increase the gas flow during the experiment due to a small leak 
in the RTPC, which concurred to the large shift of drift time observed around run number 
61600. While our gas system was kept slightly over 
atmospheric pressure to avoid contamination from air or other external gases, it 
is likely that this leak was the source of modification of the gas mixture.}.  
Figure~\ref{fig:Drift_run_number_1} shows the $T_{Max/2}$ values for 
individual runs (approximately 2 hours long). We observe a significant 
variation of the drift speed over time that need to be accounted for 
in track reconstruction.

@@

In summary, we obtain from our calibration a parame\-tri\-zation of the drift speed 
as a function of both position along the beam axis and the run number. These 
functions were extracted for our entire data set and implemented in the 
track reconstructions code.

@@

The drift path is the trajectory followed by the electrons released through 
ionization in the gas. We calculate them with
MAGBOLTZ~\cite{MAGBOLTZ}, but it requires knowledge of the detector's 
geometry, gas mixture composition, and of course the electric and magnetic 
fields over the whole volume of the detector. We used such calculation as a 
first calibration, but, 
as can be seen with the drift speed, experimental conditions were unstable in the 
chamber. Moreover, the 4 $\mu$m foil used as a cathode is easily deformed, such
that we expect the geometry accuracy to be of few millimeters, directly impacting 
our knowledge of the electric field. These problems, already 
encountered for the BoNuS RTPC calibration~\cite{BONUS-NIM}, motivated the 
acquisition of specific calibration runs. These were taken with a lower energy 
electron beam (1.204 and 1.269 GeV) to enhance the cross section of the elastic 
scattering ($e^{4}$He$\rightarrow e^{4}$He). In this process, the measurement of
the electron kinematics allows to calculate directly the kinematics of the Helium nucleus. 
By comparing the calculated momentum and angle of the recoil alpha particle to the 
measurement in the RTPC, we tuned the drift paths independently of our 
knowledge of the chamber's conditions.

Based on the kinematics of the electrons in the calibration data, 
we generate the helium nucleus in our RTPC GEANT4 simulation~\cite{GEANT4}. Then 
we compare the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber. 
Because of the magnetic field, the drift paths are not linear in the RTPC. So 
to perform the extraction, we make a first approximation with a linear 
dependence between the radius of emission and the time of detection, and then 
refine our result. As it happens, the curvature is minimal and this process 
converges already on the second iteration.

@@

To calibrate the gains, we compare the experimental ADCs to the energy deposited 
for each pad individually in GEANT4 by similar simulated tracks (using the 
same elastic events as for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the DAQ features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as shown in 
Fig.~\ref{fig:EVENT_adc_tdc}. The gain for each pad is calculated 
as the ratio of the measured ADCs to the simulated deposited energy.  
Then, these gains are refined using correction factors obtained from a 
sample of good tracks. For each track, we calculate the corrected 
energy deposit on a pad and compare it to the average
deposit recorded by the other pads. This last step was implemented to 
avoid any bias from this simulation based calibration method.
Final results are shown in Fig.~\ref{fig:dedx_p_exp_2nd}, where energy 
loss of particles is plotted against momentum over charge ration. One can 
clearly see the band for $^4$He in its expected position.

@@

\section{Track Reconstruction}
\label{sec_tracking}

@@

The first is an oscillatory noise located early in the readout time window, 
shown in the top panel of Fig.~\ref{fig:noise} for a particularly noisy 
channel. Its amplitude is not dissimilar to those of real tracks. About 
18\% of the readout channels exhibit large contributions from this noise 
characteristic. Due to its unique time-energy correlation for the given 
channels, the noise could be removed on an event by event and channel by 
channel basis without significant loss of good signals. The result of the
procedure is illustrated in the bottom panel of Fig.~\ref{fig:noise}.

The second noise signature was a coherent noise affecting about 25\% of the 
amplifiers boards, when simultaneous hits in most of the 16 channels of the
board were recorded. An event-based technique to identify and remove this noise was developed based 
on counting simultaneous hits in preamp group, and, if sufficiently large, 
perform a dynamic pedestal subtraction based on the average ADC of the 
channels' neighbors within the preamp group.

@@

\subsection{Resolution}
Elastic scattering was used to estimate the tracking resolution of the RTPC 
based on the residual between the expected and measured $^4$He tracks.  The RTPC
resolutions, after removing contributions from the electron, are shown in Table
~\ref{tab:reso}, and are very similar for the two halves of the RTPC.  Note, the
$\theta$- and $z$-resolutions are highly correlated.
\begin{table}[htbp]
\begin{center}
\begin{tabular}{|l|cccc|}
  \hline
& $\sigma_{z}$ &  $\sigma_{\theta}$ & $\sigma_{\phi}$ & $\sigma_{p}$\\
\hline
Left &  5.3 mm & 3.8$^{\circ}$ & 1.9$^{\circ}$ & 9$\%$ \\
Right & 6.5 mm & 4.0$^{\circ}$ & 1.9$^{\circ}$ & 8$\%$\\
\hline
\end{tabular}
\caption{The resolutions of the two modules of the RTPC for $z$-vertex, polar and azimuthal angles, and momentum.}
\label{tab:reso}
\end{center}
\end{table}

\subsection{Efficiency}
We measured the efficiency of the RTPC
using elastic scattering on $^4$He by comparing the inclusive yield, based
only on electron detection, to the exclusive elastic yield, where the Helium
recoil is also detected (see Fig.~\ref{fig:w}). We present in Fig.~\ref{fig:rtpc_eff}
the results for the two halves of the detector. We observe that the left and 
the right modules have similar efficiencies except near the upstream target 
window. This difference is due to the large number of dead channels in this 
area of the detector, which are concentrated on the left.
